---
title: "Compare MODIS Grids"
author: "Johnathan Rush"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results='hold')
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(cache.path = "/home/rushj03/dev/docs/cache/")
```

This document compares three MODIS products we're interested in using and tests whether their coordinate system, resolution, and registration are the same.  

# Libraries

```{r libraries, message=FALSE}
library(raster)
library(rgdal)
library(sp)
library(gdalUtils)
```

# Daily MAIAC: MCD19A2

MAIAC from NASA DAAC  

We have a four files downloaded for a test region in NEMIA: two tiles for two days.  
We'll use day 292, which is 10-19-2015, and tile h11v05. We'll select matching tiles for the other datasets.    

```{r mcd19}
cat("MCD19A2 available files:\n")
list.files("~/qnap_geo/MCD19A2/", pattern = ".*hdf$")
cat("\nSelected MCD19A2 file:\n")
list.files("~/qnap_geo/MCD19A2/", pattern = "MCD19A2.A2015292.h11v05.*hdf$")
mcd19file = list.files("~/qnap_geo/MCD19A2/", pattern = "MCD19A2.A2015292.h11v05.*hdf$", full.names = TRUE)
```

# Daily LST: MOD11A1 (Terra)

Collection 6 LST  

```{r mod11_c6}
input_path = "/data-belle/LST/MODIS.LST.C6/HDF/MOLT/MOD11A1.006/" # for terra
this_date = "2015.10.19"
cat("Selected MOD11A1 file:\n")
list.files(paste0(input_path, this_date), pattern = ".*h11v05.*hdf$")
mod11file = list.files(paste0(input_path, this_date), pattern = ".*h11v05.*hdf$", full.names = TRUE)
```

# Monthly NDVI: MOD13A3 (Terra)

Collection 5 NDVI  

```{r mod13}
input_path = "/data-belle/LST/MODIS.LST.18122016/modis_lst_ndvi_2000-2016/e4ftl01.cr.usgs.gov/MODIS_Composites_NDVI/MOLT/MOD13A3.005/" # for terra
this_date = "2015.10.01" # monthly files
cat("Selected MOD13A3 file:\n")
list.files(paste0(input_path, this_date), pattern = ".*h11v05.*hdf$")
mod13file = list.files(paste0(input_path, this_date), pattern = ".*h11v05.*hdf$", full.names = TRUE)
```

# Open one tile for each

```{r mod11_raster}
h11v05_mod11A1 = raster(get_subdatasets(mod11file)[1]) # sd 1 is LST
h11v05_mod11A1
```

```{r mcd19_raster}
h11v05_mcd19A2 = raster(get_subdatasets(mcd19file)[1]) # sd 1 is Optical_Depth_047
h11v05_mcd19A2
```

```{r mod13_raster}
h11v05_mod13A3 = raster(get_subdatasets(mod13file)[1]) # sd 1 is 1_km_monthly_NDVI
h11v05_mod13A3
```

The proj4strings are identical for all three.  

# Compare Rasters

The `raster::compareRaster` function allows us to compare the extent, dimensions, CRS, resolution, origin, and rotation of two different rasters. These are compared within a tolerance expressed as fractions of a grid cell. The default value is 1/10 of the grid cell size.  

We'll compare each of these grid attributes while shrinking the tolerance, and report when it fails.  

```{r gridComp_function}
# assumes meters, which is correct for the MODIS CRS
gridComp <- function(grid1, grid2){
  for(i in 1:9){
    rasterOptions(tolerance = 1*10^(-i))
    comp = compareRaster(grid1, grid2, res = T, orig = T, stopiffalse = F, showwarning = T)
    if(comp == FALSE) break
  }
  if(i > 1){
    #cat(paste0("\nRasters were the same at the 1e-", i-1, " level, but not at the 1e-", i, " level\n"))
    cat(paste0("\nRasters were the same within a tolerance of ", 
               format(1*10^(-i+1)*res(grid1)[[1]],scientific = T, digits = 3), 
               " meters, but not within ", 
               format(1*10^(-i)*res(grid1)[[1]],scientific = T, digits = 3), " meters\n"))
  }else{
    cat(paste0("Rasters were not the within a tolerance of ", 
               format(1*10^(-i)*res(grid1)[[1]],scientific = T, digits = 3), " meters\n"))
  }
}
```

## Compare MOD11A1 to MCD19

Terra Daily LST collection 6 vs. MAIAC Daily  

```{r compare_lst_maiac}
gridComp(h11v05_mod11A1, h11v05_mcd19A2)
```

## Compare MOD11A1 to MOD13A3

Terra Daily LST **collection 6** vs. Terra Monthly NDVI **collection 5**  

We didn't have any collection 6 NDVI downloaded for this region.  

```{r compare_lst_ndvi}
gridComp(h11v05_mod11A1, h11v05_mod13A3)
```

# Summary

These grids have *functionally identical* registrations and resolutions, and *identical* CRS and dimensions.  

## Centroid point distance

We can go even farther and check the distance between the corresponding centroids of two different rasters.  

First, we'll reassign the values to all 0 or 1. 1 will be used sparingly, and only those cells will be selected when converting to points.  

```{r raster_binary_vector}
# build vector of new values
newv = rep.int(0, ncell(h11v05_mod11A1))
skipnumber = 110
for(i in 1:as.integer(ncell(h11v05_mod11A1)/skipnumber)){
  newv[i*skipnumber] = 1
}
```

Set values of rasters
```{r reset_raster_values}
h11v05_mcd19A2[] <- newv
h11v05_mod11A1[] <- newv
names(h11v05_mcd19A2) <- "mcd19"
names(h11v05_mod11A1) <- "mod11"
```

Convert to points
```{r raster_to_points}
# takes about half a second each
spdf_mcd19_h11v05 <- rasterToPoints(h11v05_mcd19A2, spatial = TRUE, fun=function(x){x==1})
spdf_mod11_h11v05 <- rasterToPoints(h11v05_mod11A1, spatial = TRUE, fun=function(x){x==1})
paste("Same number of points for each:", nrow(spdf_mcd19_h11v05) == nrow(spdf_mod11_h11v05))
```

```{r calculate_distance}
distvec = pointDistance(spdf_mcd19_h11v05, spdf_mod11_h11v05, lonlat = FALSE)
cat(paste("Range in raster centroid distances:", paste(round(range(distvec),6), collapse = " to "), "meters\n"))
```

